The Galactic Delay-Time Distribution of Type Ia Supernovae:\\ A Chemical Evolution Perspective

### 2023-11-15 src/tex/ms.tex
[showyourwork] overleaf sync

% Define document class
\documentclass[modern,linenumbers]{aastex631}
%twocolappendix

\usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
% \usepackage{layouts}
\usepackage{xcolor}
\usepackage{upgreek}
\let\tablenum\relax
\usepackage{siunitx}

% user-defined commands
\newcommand{\yes}{\textcolor{green}{\checkmark}}
\newcommand{\meh}{\textcolor{black}{$\sim$}}
\newcommand{\no}{\textcolor{red}{$\times$}}
\newcommand{\osuaffil}{Department of Astronomy, The Ohio State University, 140 W. 18th Ave, Columbus OH 43210, USA}

@@

% Abstract from AAS 241
\begin{abstract}
    Type Ia supernovae (SNe Ia) produce most of Fe-peak elements in the Universe and therefore are a crucial ingredient in galactic chemical evolution models. SNe Ia do not explode immediately after star formation and the delay-time distribution (DTD) has not been definitively determined by supernova surveys or theoretical models. Because the DTD also affects the relationship among age, [Fe/H], and [$\alpha$/Fe] in chemical evolution models, comparison with observations of stars in the Milky Way is an important consistency check for any proposed DTD. We implement several popular forms of the DTD in combination with multiple star formation histories for the Milky Way in one-zone and multi-zone chemical evolution models using the Versatile Integrator for Chemical Evolution (VICE). VICE includes a prescription for radial stellar migration in multi-zone models. We compare our predicted interstellar medium abundance tracks, stellar abundance distributions, and stellar age distributions to the 17th data release of the Apache Point Observatory Galactic Evolution Experiment (APOGEE). We find that the SN Ia DTD has the largest effect on the [$\alpha$/Fe] distribution: a DTD with more prompt SNe Ia produces a stellar abundance distribution that is skewed toward a lower [$\alpha$/Fe] ratio, whereas a DTD with more delayed SNe Ia produces a distribution that is skewed toward higher [$\alpha$/Fe]. While the DTD alone cannot explain the observed bimodality in the [$\alpha$/Fe] distribution, in combination with an appropriate star formation history it affects the goodness of fit between the simulated and observed high-$\alpha$ sequence.
\end{abstract}

@@

[Hook and intro paragraph]
Galactic chemical evolution (GCE) studies seek to reproduce the observed distribution of metals throughout the Milky Way Galaxy.

[Introduction to SNe Ia]
Type Ia supernovae (SNe Ia), the explosions of carbon-oxygen white dwarfs (WDs), are responsible for roughly half of the iron produced in the Galaxy. Despite their importance, the mechanism(s) for producing SNe Ia are not fully understood. Two general production channels have been proposed: in the single-degenerate (SD) case, the WD accretes mass from a close non-degenerate companion until it surpasses a mass of $\sim1.4$ M$_\odot$ and explodes. In the double-degenerate (DD) case, two WDs merge after a gravitational-wave inspiral or head-on collision. Recent observations have placed tight constraints on the SD channel, heavily disfavoring it as the main pathway for producing ``normal'' SNe Ia. Though it is now the preferred model, the DD channel faces issues of its own with matching observed SN Ia rates as well as the fact that not all WD mergers may lead to a thermonuclear explosion.

[What is the delay time distribution]
The delay-time distribution (DTD) is the rate of SNe Ia which explode as a function of time after a hypothetical single star formation event. By convolving the DTD with the star formation rate (SFR) as a function of time, one can produce a function describing the time-variation in the SN Ia rate.

@@

Studies of galactic chemical evolution typically assume a single form for the SN Ia DTD. This is the first comprehensive look at the DTD in a multi-zone GCE model which includes the effects of radial migration.

@@

One curious chemical feature of the Milky Way disk is the existence of two populations separated by their [$\alpha$/Fe] abundance ratios. First noted by \citet{Furhmann1998-NearbyStars}, expanded by \citet{Hayden2015-ChemicalCartography}, definitely others too. The ``high-$\alpha$'' population is associated with the thick disk and contains stars with low enrichment from SNe Ia, so their abundance ratios reflect that of pure CCSN enrichment. The ``low-$\alpha$'' population is associated with the thin disk and contains stars with roughly solar $\alpha$-abundances.
The origin of the [$\alpha$/Fe] bimodality is debated. \citet{Schonrich2009-RadialMixing} argue that it arises due to radial migration (churning) even with a smooth star formation rate. \citet{Johnson2021-Migration} use a smooth SFR and a prescription for radial migration but are unable to reproduce the bimodality. \citet{Spitoni2021-TwoInfall} (definitely earlier papers too) argue that two distinct infall episodes can explain the bimodality. What separates the two sequences is SN Ia enrichment, so the DTD ought to have some impact on the [$\alpha$/Fe] bimodality, perhaps even determining whether or not it exists.

@@

\begin{deluxetable*}{lll}
\tablecaption{Summary of SN Ia DTD models explored in this paper.\label{tab:dtds}}
\tablehead{
\colhead{Model} & \colhead{Parameters} & \colhead{Similar to}
}
\startdata
Power-law   & $\alpha=-1.1$                 & \citet[][field]{Maoz2017-CosmicDTD}; 
                                              \citet{Wiseman2021-DESRates}              \\
Power-law   & $\alpha=-1.4$                 & \citet[][cluster]{Maoz2017-CosmicDTD}; 
                                              \citet{Heringer2019-FieldGalaxyDTD}       \\
Plateau     & $W=0.3$ Gyr, $\alpha=-1.1$    & \citet[][CLOSE DD]{Greggio2005-AnalyticalRates} \\
Plateau     & $W=1.0$ Gyr, $\alpha=-1.1$    & \citet[][WIDE DD]{Greggio2005-AnalyticalRates} \\
Triple-system   & $f_{\rm init}=0.05f_{\rm peak}$, $t_{\rm rise}=0.5$ Gyr, & \citet{Rajamuthukumar2022-TripleEvolution} \\
            & $W=0.5$ Gyr, $\alpha=-1.1$ & \\
Exponential & $\tau=1.5$ Gyr    & \citet[][SD]{Greggio2005-AnalyticalRates};
                                  \citet{Schonrich2009-RadialMixing};       \\
            &                   & \citet{Weinberg2017-ChemicalEquilibrium}  \\
Exponential & $\tau=3.0$ Gyr    & --- \\
Two-population  & $t_{\rm max}=0.05$ Gyr, $\sigma=0.015$ Gyr, & \citet{Mannucci2006-TwoPopulations} \\
            & $\tau=3.0$ Gyr & \\
\enddata
\end{deluxetable*}

@@

\paragraph{Power-law} A simple power law with slope $\alpha$:
\begin{equation}
    R_{\rm Ia}(t) \propto (t/1\,\rm{Gyr})^\alpha
    \label{eq:powerlaw-dtd}
\end{equation}
% where $A_\alpha = (\alpha+1)(t_{\rm max}^{\alpha+1} - t_{\rm min}^{\alpha+1})^{-1}$ is the normalization coefficient.
A declining power-law with $\alpha\sim-1$ is a standard assumption in many observational studies of the DTD as well as GCE simulations. \citet{Maoz2017-CosmicDTD} obtain a DTD with $\alpha=-1.07\pm0.09$ based on volumetric rates and an assumed cosmic SFH for field galaxies in redshift range $0\leq z\leq 2.25$. \citet{Wiseman2021-DESRates} obtain a similar slope of $\alpha=-1.13\pm0.05$ for field galaxies in the redshift range $0.2<z<0.6$. For galaxy clusters, \citet{Maoz2017-CosmicDTD} find a steeper DTD slope of $\alpha=-1.39^{+0.32}_{-0.05}$. \citet{Heringer2019-FieldGalaxyDTD} use a SFH-independent method to constrain the DTD for field galaxies within $0.01<z<0.2$ and find a steeper slope of $\alpha=-1.34^{+0.19}_{-0.17}$.
In this paper, we investigate the cases $\alpha=-1.1$ and $\alpha=-1.4$.

@@

\paragraph{Triple-system} A DTD based on simulations of triple-system evolution by \citet{Rajamuthukumar2022-TripleEvolution}. We approximate their DTD as a special case of the plateau model (Equation \ref{eq:plateau-dtd}) where the initial rate is quite low until an instantaneous rise to the plateau value at time $t_{\rm rise}$:
\begin{equation}
    R_{\rm Ia}(t) \propto
    \begin{cases}
        \epsilon, & t < t_{\rm rise} \\
        1, & t_{\rm rise} \leq t < W \\
        (t / W) ^ \alpha, & t \geq W
    \end{cases}
    \label{eq:triple-dtd}
\end{equation}
We take $t_{\rm rise}=0.5$ Gyr, $W=0.5$ Gyr, $\alpha=-1.1$, and $\epsilon=0.05$ so the initial rate is 5\% of the peak rate.

\paragraph{Exponential} An exponentially declining DTD with timescale $\tau$:
\begin{equation}
    R_{\rm Ia}(t) \propto \frac{1}{\tau} e^{-t/\tau}
    \label{eq:exponential-dtd}
\end{equation}
We investigate the case where $\tau=1.5$ Gyr, which has been used in previous GCE studies \citep[e.g.,][]{Schonrich2009-RadialMixing,Weinberg2017-ChemicalEquilibrium} and is similar to the analytical single-degenerate DTD from \citet{Greggio2005-AnalyticalRates}. We also investigate a longer timescale of $\tau=3$ Gyr. \citet{Matteucci1986-SupernovaEnrichment}, \citet{Stolger2020-ExponentialDTD}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.32\linewidth]{figures/onezone_powerlaw_slope.pdf}
    \includegraphics[width=0.32\linewidth]{figures/onezone_exponential_timescale.pdf}
    \includegraphics[width=0.32\linewidth]{figures/onezone_plateau_width.pdf}
    \caption{\textit{Left:} Abundance tracks in the [Fe/H]--[O/Fe] plane for one-zone chemical evolution models models which assume a power-law DTD. The dotted, dashed, and solid curves represent models with power-law slopes $\alpha=-0.8$, $-1.1$, and $-1.4$, respectively. For reference, the solid gray curve represents an exponential DTD with $\tau=3$ Gyr. In all cases a minimum SN Ia delay time of 40 Myr is assumed. The open symbols located along each curve mark logarithmic steps in simulation time. The marginal panels above and to the right of the central panel present the distributions of [Fe/H] and [O/Fe], respectively. For display purposes, these distributions are convolved with a Gaussian with a standard deviation of 0.02 dex.
    \textit{Middle:} Similar, comparing exponential DTDs with varying timescales. The solid, dashed, and dotted curves represent timescales $\tau=6$, 3, and 1.5 Gyr, respectively.
    \textit{Right:} Similar, comparing plateau DTDs with varying plateau widths. The solid, dashed, and dotted green curves represent widths $W=1$, 0.3, and 0.1 Gyr, respectively. All plateau DTD models assume a post-plateau power-law slope of $\alpha=-1.1$. For reference, the solid gray curve represents an exponential DTD with $\tau=3$ Gyr, and the dotted purple curve represents a power-law DTD $\alpha=-1.1$ and no plateau.}
    \label{fig:onezone-threepanel}
    \script{onezone_threepanel.py}
\end{figure*}

The left-hand panel of Figure \ref{fig:onezone-twopanel} shows that the minimum SN Ia delay time has a much stronger effect on the abundance track and MDFs of models which assume a power-law DTD than models which assume a plateau or exponential DTD, due to the much higher SN Ia rate in the power-law models at early times (see Figure \ref{fig:dtds}). Moreover, a power-law DTD with a long minimum delay may be observationally hard to distinguish from a plateau model. In Figure \ref{fig:onezone-twopanel}, the abundance track for the model with a power-law DTD and $t_D=150$ Myr (red dashed line) is similar to that of the plateau DTD with $W=0.3$ Gyr and $t_D=40$ Myr (blue solid line), and the [O/Fe] distributions are virtually identical.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.49\linewidth]{figures/onezone_minimum_delay.pdf}
    \includegraphics[width=0.49\linewidth]{figures/onezone_dtd.pdf}
    \caption{\textit{Left:} Comparison of one-zone models with minimum SN Ia delay times of 40 Myr (solid curves) and 150 Myr (dashed curves) for two DTDs: a power-law with $\alpha=-1.1$ (blue curves), and an exponential with $\tau=1.5$ Gyr (purple curves). The layout is similar to the plots in Figure \ref{fig:onezone-threepanel}.
    \textit{Right:} Comparison of one-zone models which assume all five functional forms for the DTD in this paper: triple-system (green), plateau with $W=1$ Gyr (cyan), exponential with $\tau=1.5$ Gyr (purple), power-law with $\alpha=-1.1$ (blue), and two-population (pink). A minimum SN Ia delay time of 40 Myr is assumed for all models.}
    \label{fig:onezone-twopanel}
    \script{onezone_twopanel.py}
\end{figure*}

@@

The early-burst model produces a bimodal distribution of [O/Fe] even out to large radii and close to the midplane, whereas the APOGEE distribution does not show a prominent high-$\alpha$ peak beyond $R_{\rm{Gal}}\sim11$ kpc.

Away from the midplane, there is a clear trend of a higher average stellar [O/Fe] at small $R_{\rm gal}$ and lower [O/Fe] at large $R_{\rm gal}$. However, the APOGEE data show a clear dichotomy between inner-galaxy, high-$\alpha$ stars and outer-galaxy, low-$\alpha$ stars, with few in between. In the inside-out and late-burst SFHs, the shift from high-$\alpha$ to low-$\alpha$ is much more gradual and the trough at [O/Fe]$\approx0.2$ does not appear. The two-infall SFH appears to produce \textit{three} modes, at [O/Fe]$\approx 0.0$, $0.2$, and $0.4$, with the first and last showing the most variation with $R_{\rm gal}$ at high latitudes. The early-burst SFH produces the closest match to the data by far: the low-$\alpha$ sequence away from the midplane shows increasing contribution with larger radii, with the innermost radii containing very few low-$\alpha$ stars. However, the high-$\alpha$ sequence contains substantial contribution from stars at large radii, which is not seen in the data.

@@

Figure \ref{fig:ofe-df-dtd} shows [O/Fe] distributions from simulations with the same SFH but a variety of different DTDs. We choose the early-burst SFH because it produces distinct high- and low-$\alpha$ sequences, which makes the effect of the DTD on the former more apparent. The most apparent effect of the DTD is to shift the mode of the high-$\alpha$ sequence. The two-population DTD, which has the most prompt SNe Ia of any model in this paper, places the high-$\alpha$ sequence at [O/Fe]$\approx 0.2$, while the triple-system DTD, which has the fewest prompt SNe Ia, places $\sim0.2$ dex higher at [O/Fe]$\approx0.4$.

@@

\begin{acknowledgments}
    Funding for the Sloan Digital Sky 
    Survey IV has been provided by the 
    Alfred P. Sloan Foundation, the U.S. 
    Department of Energy Office of 
    Science, and the Participating 
    Institutions.

@@

>>>>

% Define document class
\documentclass[modern,linenumbers]{aastex631}
% \documentclass[twocolumn,twocolappendix,linenumbers]{aastex631}
%twocolappendix

\usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
% \usepackage{layouts}
\usepackage{xcolor}
\usepackage{upgreek}
\let\tablenum\relax
\usepackage{siunitx}
\usepackage{xspace}

% user-defined commands
\newcommand{\yes}{\textcolor{green}{\checkmark}}
\newcommand{\meh}{\textcolor{black}{$\sim$}}
\newcommand{\no}{\textcolor{red}{$\times$}}
\newcommand{\osuaffil}{Department of Astronomy, The Ohio State University, 140 W. 18th Ave, Columbus OH 43210, USA}
\newcommand{\aFe}{[$\alpha$/Fe]\xspace}

@@

% Abstract from AAS 241
\begin{abstract}
    Type Ia supernovae (SNe Ia) produce most of the Fe-peak elements in the Universe and therefore are a crucial ingredient in galactic chemical evolution models. SNe Ia do not explode immediately after star formation and the delay-time distribution (DTD) has not been definitively determined by supernova surveys or theoretical models. Because the DTD also affects the relationship among age, [Fe/H], and \aFe in chemical evolution models, comparison with observations of stars in the Milky Way is an important consistency check for any proposed DTD. We implement several popular forms of the DTD in combination with multiple star formation histories for the Milky Way in one-zone and multi-zone chemical evolution models using the Versatile Integrator for Chemical Evolution (VICE). VICE includes a prescription for radial stellar migration in multi-zone models. We compare our predicted interstellar medium abundance tracks, stellar abundance distributions, and stellar age distributions to the 17th data release of the Apache Point Observatory Galactic Evolution Experiment (APOGEE). We find that the SN Ia DTD has the largest effect on the \aFe distribution: a DTD with more prompt SNe Ia produces a stellar abundance distribution that is skewed toward a lower \aFe ratio, whereas a DTD with more delayed SNe Ia produces a distribution that is skewed toward higher \aFe. While the DTD alone cannot explain the observed bimodality in the \aFe distribution, in combination with an appropriate star formation history it affects the goodness of fit between the simulated and observed high-$\alpha$ sequence. Our model results favor a DTD with longer delay times and fewer prompt SNe Ia than the fiducial $t^{-1}$ power law.
\end{abstract}

@@

% [Hook and intro paragraph]
Galactic chemical evolution studies seek to reproduce the observed distribution of metals as a function of age throughout the Milky Way Galaxy. \citet{Tinsley1979-StellarLifetimes} proposed that the non-Solar \aFe ratios seen by, e.g., \citet{Wallerstein1962-GDwarfAbundances} were caused by different stellar lifetimes for the contributors of the Fe-peak elements than the $\alpha$-elements. Type Ia supernovae (SNe Ia), the explosions of carbon-oxygen white dwarfs (WDs), are responsible for roughly half of the iron produced in the Galaxy [cite]. SNe Ia have long been known to occur in elliptical galaxies, suggesting that their progenitors may have lifetimes of up to many Gyr. \citet{Tinsley1979-StellarLifetimes} assumed a range of $\sim 3\times 10^7 - 10^8$ years for the lifetimes of SN Ia progenitors, although they argued against the hypothesis that SNe Ia are exploding white dwarfs. \citet{Matteucci1986-SupernovaEnrichment} modeled SNe Ia as the end product of binary evolution and arrived at much longer lifetimes for the progenitors, finding that SNe Ia did not contribute substantially to chemical enrichment for $\sim 10^8$ years after the commencement of star formation. 
% [Concluding sentence about how understanding the DTD is important for chemical evolution and for understanding the progenitors of SNe Ia]
The timescale for Fe enrichment in the Galaxy, a key component of chemical evolution models, is directly linked to the progenitor lifetimes of SNe Ia.

% [What is the delay time distribution]
The delay-time distribution (DTD) is the rate of SNe Ia which explode as a function of time after a hypothetical instantaneous star formation event. By convolving the DTD with the star formation rate (SFR) as a function of time, one can produce a function describing the time-variation in the SN Ia rate.
Different models for the SN Ia progenitor have different DTDs.

% [Introduction to SNe Ia]
% Type Ia supernovae (SNe Ia), the explosions of carbon-oxygen white dwarfs (WDs), are responsible for roughly half of the iron produced in the Galaxy [cite]. Despite their importance, 
The mechanism(s) for producing SNe Ia are not fully understood \citep[for reviews, see][]{Maoz2014-Review,Livio2018-ProgenitorReview,Ruiter2020-ProgenitorReview}. Two general production channels have been proposed: in the single-degenerate (SD) case, the WD accretes mass from a close non-degenerate companion until it surpasses a mass of $\sim1.4$ M$_\odot$ and explodes \citep{Whelan1973-SDModel,Nomoto1982-SDModel,Yoon2003-SDModel}. In the double-degenerate (DD) case, two WDs merge after a gravitational-wave inspiral \citep{Iben1984-IaBinary,Webbink1984-DDModel,Pakmor2012-WDMerger} or head-on collision \citep{Benz1989-CollisionalDD,Thompson2011-CollisionalDD}. Recent observations have placed tight constraints on the SD channel, heavily disfavoring it as the main pathway for producing ``normal'' SNe Ia \citep[e.g.,][]{Panagia2006-RadioEmission,Schaefer2012-ExCompanionSNR,Chomiuk2016-RadioEmission,Fausnaugh2019-EarlyIaLightCurves,Tucker2020-SNeIaSpectra,Do2021-Progenitor1972E,Tucker2023-SN2011fe}. Though it is now the preferred model, the DD channel faces issues of its own with matching observed SN Ia rates because not all WD mergers may lead to a thermonuclear explosion \citep[e.g.,][]{Shen2012-DDMergers}.

@@

Studies of galactic chemical evolution typically assume a single form for the SN Ia DTD. This is the first comprehensive look at the DTD in a multi-zone chemical evolution model which includes the effects of radial migration.

@@

One curious chemical feature of the Milky Way disk is the existence of two populations separated by their \aFe abundance ratios at the same metallicity. First noted by \citet{Furhmann1998-NearbyStars}, expanded by \citet{Hayden2015-ChemicalCartography}, definitely others too. The ``high-$\alpha$'' population contains stars with low enrichment from SNe Ia, so their abundance ratios reflect that of pure CCSN enrichment; it is associated with the kinematically hot thick disk \citep{Bensby2003-AbundanceTrends}. The ``low-$\alpha$'' population is associated with the thin disk and contains stars with roughly solar $\alpha$-abundances.
The origin of the \aFe bimodality is debated. \citet{Schonrich2009-RadialMixing} argue that it arises due to radial migration (churning) even with a smooth star formation rate. \citet{Johnson2021-Migration} use a smooth SFR and a prescription for radial migration but are unable to reproduce the bimodality. \citet{Spitoni2021-TwoInfall} (definitely earlier papers too) argue that two distinct infall episodes can explain the bimodality. What separates the two sequences is the ratio of CCSN to SN Ia enrichment, so the DTD ought to have some impact on the \aFe bimodality, perhaps even determining whether or not it exists.

@@

\begin{deluxetable*}{lll}
\tablecaption{Summary of SN Ia DTD models explored in this paper.\label{tab:dtds}}
\tablehead{
\colhead{Model} & \colhead{Parameters} & \colhead{Similar to}
}
\startdata
Power-law   & $\alpha=-1.1$                 & \citet[][field]{Maoz2017-CosmicDTD}; 
                                              \citet{Wiseman2021-DESRates}              \\
Power-law   & $\alpha=-1.4$                 & \citet[][cluster]{Maoz2017-CosmicDTD}; 
                                              \citet{Heringer2019-FieldGalaxyDTD}       \\
Plateau     & $W=0.3$ Gyr, $\alpha=-1.1$    & \citet[][CLOSE DD]{Greggio2005-AnalyticalRates} \\
Plateau     & $W=1.0$ Gyr, $\alpha=-1.1$    & \citet[][WIDE DD]{Greggio2005-AnalyticalRates} \\
Triple-system   & $f_{\rm init}=0.05f_{\rm peak}$, $t_{\rm rise}=0.5$ Gyr, & \citet{Rajamuthukumar2023-TripleEvolution} \\
            & $W=0.5$ Gyr, $\alpha=-1.1$ & \\
Exponential & $\tau=1.5$ Gyr    & \citet[][SD]{Greggio2005-AnalyticalRates};
                                  \citet{Schonrich2009-RadialMixing};       \\
            &                   & \citet{Weinberg2017-ChemicalEquilibrium}  \\
Exponential & $\tau=3.0$ Gyr    & --- \\
Two-population  & $t_{\rm max}=0.05$ Gyr, $\sigma=0.015$ Gyr, & \citet{Mannucci2006-TwoPopulations} \\
            & $\tau=3.0$ Gyr & \\
\enddata
\end{deluxetable*}

@@

\paragraph{Power-law} A simple power law with slope $\alpha$:
\begin{equation}
    R_{\rm Ia}(t) \propto (t/1\,\rm{Gyr})^\alpha
    \label{eq:powerlaw-dtd}
\end{equation}
% where $A_\alpha = (\alpha+1)(t_{\rm max}^{\alpha+1} - t_{\rm min}^{\alpha+1})^{-1}$ is the normalization coefficient.
A declining power-law with $\alpha\sim-1$ is a standard assumption in many observational studies of the DTD as well as Galactic chemical evolution simulations. \citet{Maoz2017-CosmicDTD} obtain a DTD with $\alpha=-1.07\pm0.09$ based on volumetric rates and an assumed cosmic SFH for field galaxies in redshift range $0\leq z\leq 2.25$. \citet{Wiseman2021-DESRates} obtain a similar slope of $\alpha=-1.13\pm0.05$ for field galaxies in the redshift range $0.2<z<0.6$. For galaxy clusters, \citet{Maoz2017-CosmicDTD} find a steeper DTD slope of $\alpha=-1.39^{+0.32}_{-0.05}$. \citet{Heringer2019-FieldGalaxyDTD} use a SFH-independent method to constrain the DTD for field galaxies within $0.01<z<0.2$ and find a steeper slope of $\alpha=-1.34^{+0.19}_{-0.17}$.
In this paper, we investigate the cases $\alpha=-1.1$ and $\alpha=-1.4$.

@@

\paragraph{Triple-system} A DTD based on simulations of triple-system evolution by \citet{Rajamuthukumar2023-TripleEvolution}. We approximate their DTD as a special case of the plateau model (Equation \ref{eq:plateau-dtd}) where the initial rate is quite low until an instantaneous rise to the plateau value at time $t_{\rm rise}$:
\begin{equation}
    R_{\rm Ia}(t) \propto
    \begin{cases}
        \epsilon, & t < t_{\rm rise} \\
        1, & t_{\rm rise} \leq t < W \\
        (t / W) ^ \alpha, & t \geq W
    \end{cases}
    \label{eq:triple-dtd}
\end{equation}
We take $t_{\rm rise}=0.5$ Gyr, $W=0.5$ Gyr, $\alpha=-1.1$, and $\epsilon=0.05$ so the initial rate is 5\% of the peak rate.

\paragraph{Exponential} An exponentially declining DTD with timescale $\tau$:
\begin{equation}
    R_{\rm Ia}(t) \propto \frac{1}{\tau} e^{-t/\tau}
    \label{eq:exponential-dtd}
\end{equation}
We investigate the case where $\tau=1.5$ Gyr, which has been used in previous chemical evolution studies \citep[e.g.,][]{Schonrich2009-RadialMixing,Weinberg2017-ChemicalEquilibrium} and is similar to the analytical single-degenerate DTD from \citet{Greggio2005-AnalyticalRates}. We also investigate a longer timescale of $\tau=3$ Gyr. \citet{Matteucci1986-SupernovaEnrichment}, \citet{Stolger2020-ExponentialDTD}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/onezone_threepanel.pdf}
    % \includegraphics[width=0.32\linewidth]{figures/onezone_powerlaw_slope.pdf}
    % \includegraphics[width=0.32\linewidth]{figures/onezone_exponential_timescale.pdf}
    % \includegraphics[width=0.32\linewidth]{figures/onezone_plateau_width.pdf}
    \caption{\textit{Left:} Abundance tracks in the [Fe/H]--[O/Fe] plane for one-zone chemical evolution models models which assume a power-law DTD. The dotted, dashed, and solid curves represent models with power-law slopes $\alpha=-0.8$, $-1.1$, and $-1.4$, respectively. For reference, the solid gray curve represents an exponential DTD with $\tau=3$ Gyr. In all cases a minimum SN Ia delay time of 40 Myr is assumed. The open symbols located along each curve mark logarithmic steps in simulation time. The marginal panels above and to the right of the central panel present the distributions of [Fe/H] and [O/Fe], respectively. For display purposes, these distributions are convolved with a Gaussian with a standard deviation of 0.02 dex.
    \textit{Middle:} Similar, comparing exponential DTDs with varying timescales. The solid, dashed, and dotted curves represent timescales $\tau=6$, 3, and 1.5 Gyr, respectively.
    \textit{Right:} Similar, comparing plateau DTDs with varying plateau widths. The solid, dashed, and dotted green curves represent widths $W=1$, 0.3, and 0.1 Gyr, respectively. All plateau DTD models assume a post-plateau power-law slope of $\alpha=-1.1$. For reference, the solid gray curve represents an exponential DTD with $\tau=3$ Gyr, and the dotted purple curve represents a power-law DTD $\alpha=-1.1$ and no plateau.}
    \label{fig:onezone-threepanel}
    \script{onezone_threepanel.py}
\end{figure*}

One parameter of the DTD which we hold fixed in the multi-zone models is the minimum SN Ia delay time $t_D$. Typical choices for $t_D$ in the literature range from $\sim 40$ Myr to 150 Myr. The left-hand panel of Figure \ref{fig:onezone-twopanel} shows that the minimum SN Ia delay time has a much stronger effect on the abundance track and MDFs of models which assume a power-law DTD than models which assume a plateau or exponential DTD, due to the much higher SN Ia rate in the power-law models at early times (see Figure \ref{fig:dtds}). Moreover, a power-law DTD with a long minimum delay may be observationally hard to distinguish from a plateau model. In Figure \ref{fig:onezone-twopanel}, the abundance track for the model with a power-law DTD and $t_D=150$ Myr (dashed purple line) is similar to that of the plateau DTD with $W=0.3$ Gyr and $t_D=40$ Myr (solid green line), and the [O/Fe] distributions are virtually identical.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/onezone_twopanel.pdf}
    % \includegraphics[width=0.49\linewidth]{figures/onezone_minimum_delay.pdf}
    % \includegraphics[width=0.49\linewidth]{figures/onezone_dtd.pdf}
    \caption{\textit{Left:} Comparison of one-zone models with minimum SN Ia delay times of 40 Myr (solid curves) and 150 Myr (dashed curves) for two DTDs: a power-law with $\alpha=-1.1$ (blue curves), and an exponential with $\tau=1.5$ Gyr (purple curves). The layout is similar to the plots in Figure \ref{fig:onezone-threepanel}. For clarity, we use a mass-loading factor $\eta=1$ for the exponential DTD run.
    \textit{Right:} Comparison of one-zone models which assume all five functional forms for the DTD in this paper: triple-system (green), plateau with $W=1$ Gyr (cyan), exponential with $\tau=1.5$ Gyr (purple), power-law with $\alpha=-1.1$ (blue), and two-population (pink). A minimum SN Ia delay time of 40 Myr is assumed for all models.}
    \label{fig:onezone-twopanel}
    \script{onezone_twopanel.py}
\end{figure*}

@@

The distribution of [O/Fe] serves as a record of the relative rates of SNe Ia and CCSNe. As such, the shape of the distribution is greatly affected by both the SFH and DTD. Figure \ref{fig:ofe-df-sfh} shows the distribution of [O/Fe] across the disk for the four model SFHs, all assuming an exponential DTD with $\tau=1.5$ Gyr, compared to the distributions measured by APOGEE. We see similar trends with Galactic region across all four simulations. Near the midplane, the distributions at the innermost and outermost annuli are similar, but away from the midplane, there is a clear trend of a higher average stellar [O/Fe] at small $R_{\rm gal}$ and lower [O/Fe] at large $R_{\rm gal}$.

While trends with $R_{\rm gal}$ and $|z|$ are similar, the shape of the distribution varies greatly with the chosen SFH. The inside-out and late-burst models produce similar distributions because of the similarity of their underlying SFHs \textemdash both skew heavily toward near-Solar [O/Fe], although the late-burst model produces a slightly broader peak and a less-prominent high-[O/Fe] tail. The only region which shows any significant skew toward high [O/Fe] is $R_{\rm gal}=3-5$ kpc and $|z|=1-2$ kpc, but the shift to higher [O/Fe] at high latitudes is gradual and does not produce the notable trough at [O/Fe]$\approx0.2$ which is seen in the APOGEE data.

On the other hand, the early-burst model produces a bimodal distribution of [O/Fe] even out to large radii and close to the midplane, whereas the APOGEE distribution does not show a prominent high-$\alpha$ peak beyond $R_{\rm{Gal}}\sim11$ kpc. 
The early-burst SFH produces the closest match to the data by far: the low-$\alpha$ sequence away from the midplane shows increasing contribution with larger radii, with the innermost radii containing very few low-$\alpha$ stars. However, the high-$\alpha$ sequence contains substantial contribution from stars at large radii, which is not seen in the data.

The two-infall SFH model produces \textit{three} distinct modes at [O/Fe]$\approx 0.0$, $0.2$, and $0.4$. At small $R_{\rm gal}$ and with increasing $|z|$, the low-$\alpha$ peak decreases in prominence and the high-$\alpha$ peak increases, but the intermediate peak is a striking feature at all latitudes which is at odds with the APOGEE data.

@@

Figure \ref{fig:ofe-df-dtd} shows [O/Fe] distributions produced by simulations with the same SFH but a range of different DTDs. We choose the early-burst SFH because it produces distinct high- and low-$\alpha$ sequences, which makes the effect of the DTD on the former more apparent. The most apparent effect of the DTD is to shift the mode of the high-$\alpha$ sequence. The two-population DTD, which has the most prompt SNe Ia of any model in this paper, places the high-$\alpha$ sequence at [O/Fe]$\approx 0.2$, while the triple-system DTD, which has the fewest prompt SNe Ia, places it $\sim0.2$ dex higher at [O/Fe]$\approx0.4$. This metric alone would seem to favor the exponential ($\tau=1.5$ Gyr) or plateau ($W=0.3$ Gyr; not shown) models; however,

@@

\begin{acknowledgments}
    [Personal acknowledgements]

    Funding for the Sloan Digital Sky 
    Survey IV has been provided by the 
    Alfred P. Sloan Foundation, the U.S. 
    Department of Energy Office of 
    Science, and the Participating 
    Institutions.

@@

    [Ohio State \& Carnegie land acknowledgements]

    [David and Jennifer NSF grant]
